Stateless Model Checking with Data-Race Preemption Points

### 2016-07-19 oopsla-16/evaluation.tex
a small touchup, and move the table to fix latex layout
sully revisions

To evaluate the benefits of data-race PPs and Iterative Deepening separately, we ran the test suite under \quicksand~in \revisionx{three} different configurations, each of which was given a 1-hour budget and 10 CPUs for each test.
\begin{itemize}
		\revisionx{
	\item {\bf QS-Limited-HB}: \quicksand~using Limited-HB to add data-race PPs,
	\item {\bf QS-Pure-HB}: \quicksand~using pure HB instead, and
		}
	\item {\bf QS-Sync-Only}: \quicksand~with PPs still seeded as described in \sect{\ref{sec:initial-pp}}, but never adding new PPs from reported data-race candidates.
\end{itemize}

@@

\begin{figure}[t]
	\hspace{-1em}
		\revision{
	\begin{tabular}{p{0.48\textwidth}}
	\includegraphics[width=0.48\textwidth]{dowefindbugsfaster-v2.pdf} \\
		(a) Bugs found \revisionx{as a function of} elapsed CPU time. Overall,
		a more \revisionx{resource-fair} comparison than (b),
		although \quicksand's start-up overhead is exaggerated, as the SSS-MC tests are not parallelized. \\
		\\
	\includegraphics[width=0.48\textwidth]{dowefindbugsfaster-wallclock-v2.pdf} \\
		(b) Bugs found by elapsed wall-clock time.
		\quicksand~is parallelized tenfold; the vertical line indicates its 1 hour limit. \\
	\end{tabular}
	}
	\caption{Comparison of bug-finding performance
	by several configurations of \quicksand~and the SSS-MC control.
	\quicksand~finds \revisionx{125}\% as many bugs with data-race PPs.}
	\label{fig:dowefindbugsfaster}
\end{figure}

@@

{\bf Full verification.}
For 153 of our 629 tests, \revisionx{QS-Limited-HB} was able to provide the total verification guarantee described in \sect{\ref{sec:totalverif}},
\revisionx{and QS-Pure-HB completed a verification for 167 tests.}
In Figure~\ref{fig:totalverif} we plot the cumulative distribution of
%total
verifications provided by each approach.
\revisionx{The next best approach for verifications was SSS-MC-Shared-Mem, which completed its search in only 39 tests.}

>>>>

To evaluate the benefits of data-race PPs and Iterative Deepening separately, we ran the test suite under \quicksand~in \revisionx{three} different configurations, each of which was given a 1-hour budget and 10 CPUs for each test.
\begin{itemize}
		\revisionx{
	\item {\bf QS-Limited-HB}: \quicksand~using Limited-HB for its data-race analysis,
	\item {\bf QS-Pure-HB}: \quicksand~using pure HB instead, and
		}
	\item {\bf QS-Sync-Only}: \quicksand~with PPs still seeded as described in \sect{\ref{sec:initial-pp}}, but never adding new PPs from reported data-race candidates.
\end{itemize}

@@

\begin{figure}[t]
	\hspace{-1em}
		\revision{
	\begin{tabular}{p{0.48\textwidth}}
	\includegraphics[width=0.48\textwidth]{dowefindbugsfaster-v2.pdf} \\
		(a) Bugs found \revisionx{as a function of} elapsed CPU time. Overall,
		a more \revisionx{resource-fair} comparison than (b),
		although \quicksand's start-up overhead is exaggerated, as the SSS-MC tests are not parallelized. \\
		\\
	\includegraphics[width=0.48\textwidth]{dowefindbugsfaster-wallclock-v2.pdf} \\
		(b) Bugs found by elapsed wall-clock time.
		\quicksand~is parallelized tenfold; the vertical line indicates its 1 hour limit. \\
	\end{tabular}
	}
	\caption{Comparison of bug-finding performance
	by several configurations of \quicksand~and the SSS-MC control.
	\quicksand~finds \revisionx{125\%} as many bugs with data-race PPs.}
	\label{fig:dowefindbugsfaster}
\end{figure}

@@

{\bf Full verification.}
For 153 of our 629 tests, \revisionx{QS-Limited-HB} was able to provide the total verification guarantee described in \sect{\ref{sec:totalverif}},
\revisionx{and QS-Pure-HB completed a verification for 167 tests.}
In Figure~\ref{fig:totalverif} we plot the cumulative distribution of
%total
verifications provided by each approach.
\revisionx{The next best approach for verifications was SSS-MC-Shared-Mem, which completed its search in only 39 cases.}

### 2016-07-19 oopsla-16/intro.tex
sully revisions

However, all stateless model checkers to date are bound by a fixed set of {\em preemption points}: code locations that define the granularity at which threads interleave.
For example, \textsc{CHESS} \cite{chess} \revisionx{by default} preempts only on synchronization operations and library calls, which can miss lock-free shared memory races.
%
On the other hand, SPIN \cite{spin} %and Inspect \cite{inspect}
%are able to preempt
preempts threads before any shared memory access.
Such fine granularity would automatically check each data race for the possibility of failure, but risks timing out before the state space can be completed. %makes full state space completion intractable for even modestly-sized tests.
\revisionx{Some tools, such as CHESS and Inspect \cite{inspect} can strike a middle ground using compiler instrumentation to statically add preemption points on memory accesses.
Nevertheless,}
choosing preemption points is a tradeoff between schedule coverage and feasibility of completion:
%but
even with state-of-the-art reduction techniques,
\revision{fixing the degree of coverage}~in advance necessarily leaves some tests unaffordably large \cite{parrot,mcr}.
%This work shows how to avoid making that tradeoff decision in advance.

@@

Our contributions are as follows:
\begin{enumerate}
	\item {\em Iterative Deepening}, a new \revision{algorithm}~for combining data-race analysis with stateless model checking, and \quicksand, an open-source implementation;
	\item A proof of convergence \revisionx{for sequentially-consistent memory models}, showing that should it be possible in the given CPU budget,
		fully testing every discovered data-race preemption point is equivalent to testing all possible thread schedules;
	\item A new tactic for eliminating one class of false-positive data race candidates,
		which cannot soundly be used in a single-pass analysis,
		but which we prove correct when used with Iterative Deepening;
	\item A large evaluation in which \quicksand~compares favourably to stand-alone data-race detection and stateless model checking approaches, finding new bugs that would be missed by either alone.
\end{enumerate}

>>>>

However, all stateless model checkers to date are bound by a fixed set of {\em preemption points}: code locations that define the granularity at which threads interleave.
For example, \textsc{CHESS} \cite{chess} \revisionx{by default} preempts only on synchronization operations and library calls, which can miss lock-free shared memory races.
%
On the other hand, SPIN \cite{spin} %and Inspect \cite{inspect}
%are able to preempt
preempts threads before any shared memory access.
Such fine granularity would automatically check each data race for the possibility of failure, but risks timing out before the state space can be completed. %makes full state space completion intractable for even modestly-sized tests.
\revisionx{Some tools, such as CHESS and Inspect \cite{inspect} can strike a middle ground by using compiler instrumentation to statically add preemption points on memory accesses.
Nevertheless,}
choosing preemption points is a tradeoff between schedule coverage and feasibility of completion:
%but
even with state-of-the-art reduction techniques,
\revision{fixing the degree of coverage}~in advance necessarily leaves some tests unaffordably large \cite{parrot,mcr}.
%This work shows how to avoid making that tradeoff decision in advance.

@@

Our contributions are as follows:
\begin{enumerate}
	\item {\em Iterative Deepening}, a new \revision{algorithm}~for combining data-race analysis with stateless model checking, and \quicksand, an open-source implementation;
	\item A proof of convergence, showing that should it be possible in the given CPU budget,
		fully testing every discovered data-race preemption point is equivalent to testing all possible thread schedules \revisionx{(assuming a sequentially-consistent memory model)};
	\item A new tactic for eliminating one class of false-positive data race candidates,
		which cannot soundly be used in a single-pass analysis,
		but which we prove correct when used with Iterative Deepening;
	\item A large evaluation in which \quicksand~compares favourably to stand-alone data-race detection and stateless model checking approaches, finding new bugs that would be missed by either alone.
\end{enumerate}

### 2016-07-19 oopsla-16/overview.tex
sully revisions

\revisionx{
{\bf Philosophy of bugs.}
% two camps: all races are bugs; citations:
% and, races must be classified; citations: replay analysis, portend,
While there is a vast body of work on how to detect data races to begin with,
judging data races once found is a matter of philosophical debate unto itself.
Some recent tools classify races depending on how they impact program behaviour \cite{recordreplaydrs,portend}, overlooking {\em benign} races in search of more dangerous ones.
%In particular,
\cite{racerx} acknowledges that a program may have too many races for a user to worry about,
so bug reports must be prioritized by severity of effects. %how likely they may cause failures or security vulnerabilities.
%
However, other prior work argues that data races are always bugs \cite{miscompile-benign,data-races-are-evil},
largely due to the possibility of compiler or hardware reordering of racy accesses.
%
We take the former camp:
we consider a data race a bug only when it results in a visible failure state (e.g., crash or deadlock).
We bypass concerns of compiler reordering by checking programs at the executable level;
for a study of hardware reordering in the context of DPOR, we refer the reader to \cite{tsopso}.
}

>>>>

\revisionx{
{\bf Philosophy of bugs.}
% two camps: all races are bugs; citations:
% and, races must be classified; citations: replay analysis, portend,
While there is a vast body of work on how to detect data races to begin with,
judging data races once found is a matter of philosophical debate unto itself.
Some recent tools classify races depending on how they impact program behaviour \cite{recordreplaydrs,portend}, overlooking {\em benign} races in search of more dangerous ones.
%In particular,
\cite{racerx} acknowledges that a program may have too many races for a user to worry about,
so bug reports must be prioritized by severity of effects. %how likely they may cause failures or security vulnerabilities.
%
However, other prior work argues that data races are always bugs \cite{miscompile-benign,data-races-are-evil},
largely due to the possibility of compiler or hardware reordering of racy accesses.
%
C++ declares any race between two non-atomic locations to be undefined behaviour \cite{cpp-foundation}.
We take the former camp:
we consider a data race a bug only when it results in a visible failure state (e.g., crash or deadlock).
(From a C++ perspective, we assume all concurrent accesses are implemented by {\tt std::atomic}.)
We bypass concerns of compiler reordering by checking programs at the executable level;
for a study of hardware reordering in the context of DPOR, we refer the reader to \cite{tsopso}.
}
